---
title: "Syntax Highlighting in RMarkdown"
author: "Curtis Alexander"
output: 
  html_document:
    css: highlight.css
    highlight: pygments
    keep_md: yes
    theme: readable
---

### Preamble
Thanks to [Yihui](https://github.com/yihui/knitr/issues/1185), I now know that RMarkdown utilizes the [highlighting-kate](https://github.com/jgm/highlighting-kate) syntax highlighter.  The list of languages that are supported can be found [here](https://github.com/jgm/highlighting-kate/tree/master/xml).

If I understand correctly, there is a three step process to take an `Rmd` file and produce an `HTML` file.  Forgive the rough pseudo code.

```{r eval=FALSE}
md <- Rmd %>% knitr
html <- md %>% rmarkdown %>% pandoc
```

### SAS Syntax Highlighting 
Unfortunately, I wanted SAS highlighting to work within RMarkdown. This would require me submitting a pull request to [highlighting-kate](https://github.com/jgm/highlighting-kate) for SAS highlighting.  Rather than creating a syntax file, I set about figuring out how to make use of `knitr` and the `highr` package to get the highlighting I was after.

Below is an example of highlighting SAS code within an RMarkdown document.  I _thought_ you could simply use the `{r engine = 'sas', eval = FALSE}` chunk options to highlight SAS code.  Unfortunately this does not work as the highlighting is being done by `pandoc` and not `knitr`.

Instead, I created a source hook to appropriately highlight and display the SAS code.  The SAS code is put through the [highlight](http://www.andre-simon.de/doku/highlight/en/highlight.php) syntax highlighting program via the [highr](https://cran.r-project.org/web/packages/highr/index.html) package.  The resulting SAS + CSS code is wrapped in the appropriate html tags using the source hook, which produces HTML within the RMarkdown document of the form below.

```
<div class="sourceCode">
  <pre class="sourceCode">
    <code class="sourceCode">
    ** SAS code + CSS goes here **
    </code>
  </pre> 
</div>
```

### Source Hook
Below is the actual source hook needed.  It is based on an example from [Ramnath Vaidyanathan](http://ramnathv.github.io/posts/verbatim-chunks-knitr/index.html) and **tons** of trial and error.

The source hook actually works with other languages as well.  See the [Alternative Syntax Highlighting](#alternative-syntax-highlighting) section below.

```{r setup}
knitr::knit_hooks$set(source = function(x, options) {
  if (!is.null(options$hilang)) {
      code_open <- "\n\n<div class=\"sourceCode\">\n<pre class=\"sourceCode\">\n<code class=\"sourceCode\">"
      code_close <- "\n</code>\n</pre>\n</div>\n"
      code_body <- highr::hi_andre(x, language = options$hilang, format = "html")
    stringr::str_c(
      code_open,
      knitr:::indent_block(paste(code_body, collapse = '\n'), ""),
      code_close
    )
  } else {
    stringr::str_c("\n\n```", tolower(options$engine), "\n",
                   paste(x, collapse = '\n'), "\n```\n\n")
      
  }
})
```

When you want to actually highlight SAS code, you need utilize the `eval = FALSE` and `hilang = 'sas'` options rather than utilizing the `engine = 'sas'` option.

For example, if you put the following within your RMarkdown document 

<div class="sourceCode"><pre class="sourceCode"><code class="sourceCode">```{r eval = FALSE, hilang = 'sas'}
data _null_;
  x = 1;
  /* comment */
run;
```</code></pre></div>

then the result would be the following.

```{r eval = FALSE, hilang = 'sas'}
data _null_;
  x = 1;
  /* comment */
run;
```

**Note:** in order to display the backticks and chunk options, I followed the guidance at the [RMarkdown site](http://rmarkdown.rstudio.com/articles_verbatim.html).

### Alternative Syntax Highlighting
The source hook allows for highlighting any source code that the [highlight](http://www.andre-simon.de/doku/highlight/en/highlight.php) program supports.  A full list can be found [here](http://www.andre-simon.de/doku/highlight/en/langs.php).

In order to utilize the source hook, the syntax name passed to the `hilang` chunk option must match the name needed for the `highlight` program.  A full list of specific names can be found at the webpage above or by running `highlight --list-langs` from the command line.

As another example, below is code taken from the [Go](https://golang.org/) language's home page.


```{r eval = FALSE, hilang = 'go'}
package main

import "fmt"

func main() {
	fmt.Println("Hello, 世界")
}
```

### Assumptions
* The following R packages need to be installed:
    * [highr](https://cran.r-project.org/web/packages/highr/index.html)
    * [stringr](https://cran.r-project.org/web/packages/stringr/index.html)
* The [highlight](http://www.andre-simon.de/doku/highlight/en/highlight.php) program needs to be installed.
* Use the `css: highlight.css` option within the YAML header of your RMarkdown file.  The `highlight.css` file is the CSS file produced by the `highlight` program.  It can be copied to the appropriate directory after running the following.

```
highlight --print-style
```
